CRUSE to Safe Cycling in Ireland \\\large{An Open Source Tool to Support Active Travel}

### 2023-10-09 paper.tex
Shorten abstract

\usepackage{amsmath,amssymb}
\usepackage{iftex}
\ifPDFTeX
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
  \usepackage{textcomp} % provide euro and other symbols
\else % if luatex or xetex
  \usepackage{unicode-math}
  \defaultfontfeatures{Scale=MatchLowercase}
  \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
\fi
\usepackage{lmodern}
\ifPDFTeX\else  
    % xetex/luatex font selection
\fi
% Use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\IfFileExists{microtype.sty}{% use microtype if available
  \usepackage[]{microtype}
  \UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\makeatletter
\@ifundefined{KOMAClassName}{% if non-KOMA class
  \IfFileExists{parskip.sty}{%
    \usepackage{parskip}
  }{% else
    \setlength{\parindent}{0pt}
    \setlength{\parskip}{6pt plus 2pt minus 1pt}}
}{% if KOMA class
  \KOMAoptions{parskip=half}}
\makeatother
\usepackage{xcolor}
\setlength{\emergencystretch}{3em} % prevent overfull lines
\setcounter{secnumdepth}{5}
% Make \paragraph and \subparagraph free-standing
\ifx\paragraph\undefined\else
  \let\oldparagraph\paragraph
  \renewcommand{\paragraph}[1]{\oldparagraph{#1}\mbox{}}
\fi
\ifx\subparagraph\undefined\else
  \let\oldsubparagraph\subparagraph
  \renewcommand{\subparagraph}[1]{\oldsubparagraph{#1}\mbox{}}
\fi

@@

\makeatletter
\makeatother
\makeatletter
\makeatother
\makeatletter
\@ifpackageloaded{caption}{}{\usepackage{caption}}
\AtBeginDocument{%
\ifdefined\contentsname
  \renewcommand*\contentsname{Table of contents}
\else
  \newcommand\contentsname{Table of contents}
\fi
\ifdefined\listfigurename
  \renewcommand*\listfigurename{List of Figures}
\else
  \newcommand\listfigurename{List of Figures}
\fi
\ifdefined\listtablename
  \renewcommand*\listtablename{List of Tables}
\else
  \newcommand\listtablename{List of Tables}
\fi
\ifdefined\figurename
  \renewcommand*\figurename{Figure}
\else
  \newcommand\figurename{Figure}
\fi
\ifdefined\tablename
  \renewcommand*\tablename{Table}
\else
  \newcommand\tablename{Table}
\fi
}
\@ifpackageloaded{float}{}{\usepackage{float}}
\floatstyle{ruled}
\@ifundefined{c@chapter}{\newfloat{codelisting}{h}{lop}}{\newfloat{codelisting}{h}{lop}[chapter]}
\floatname{codelisting}{Listing}
\newcommand*\listoflistings{\listof{codelisting}{List of Listings}}
\makeatother
\makeatletter
\@ifpackageloaded{caption}{}{\usepackage{caption}}
\@ifpackageloaded{subcaption}{}{\usepackage{subcaption}}
\makeatother
\makeatletter
\@ifpackageloaded{tcolorbox}{}{\usepackage[skins,breakable]{tcolorbox}}
\makeatother
\makeatletter
\@ifundefined{shadecolor}{\definecolor{shadecolor}{rgb}{.97, .97, .97}}
\makeatother
\makeatletter
\makeatother
\makeatletter
\makeatother
\journal{Journal Name}
\ifLuaTeX
  \usepackage{selnolig}  % disable illegal ligatures
\fi
\usepackage[]{natbib}
\bibliographystyle{elsarticle-num}
\IfFileExists{bookmark.sty}{\usepackage{bookmark}}{\usepackage{hyperref}}
\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\urlstyle{same} % disable monospaced font for URLs
\hypersetup{
  pdftitle={CRUSE to Safe Cycling in Ireland},
  pdfauthor={Robin Lovelace; Joey Talbot; Eugeni Vidal Tortosa; Hussein Mahfouz; Elaine Brick; Peter Wright; Gary O'Tool; Dan Brennan; Suzanne Meade},
  pdfkeywords={Cycling, Open-source, Road Safety, Active Travel},
  colorlinks=true,
  linkcolor={blue},
  filecolor={Maroon},
  citecolor={Blue},
  urlcolor={Blue},
  pdfcreator={LaTeX via pandoc}}

@@

\author[1]{Eugeni Vidal Tortosa%
%
}

@@

\begin{abstract}
Under the EU Road infrastructure safety management (RISM) directive, the
National Road Safety Strategy (RSS), and the Climate Action Plan
Transport Infrastructure Ireland (TII) has a remit for road safety and
decarbonizing a predominantly road-based network in Ireland.

The Cycle Route Uptake and Scenario Estimation (CRUSE) Tool was
commissioned by TII and developed by the Leeds Institute for Transport
Studies (ITS) and AECOM To address data needs for safety, business case
development and project evaluation on the National Road Network (NRN).
Cycling in Ireland represents only 3\% of total modal share as of the
latest Census (2016), but accounts for 20\% of serious injuries and 7\%
of all fatalities. High resolution datasets on baseline and potential
cycling levels are needed to meet Ireland's climate and safety targets.

The CRUSE project builds on the methods underlying the Propensity to
Cycle Tool for England and Wales (PCT), which had a major impact on
strategic cycle network planning in the UK. Like the PCT, CRUSE is based
on open source software and reproducible data science. The resulting web
application is available at \url{https://cruse.bike/}, enabling
planners, engineers, and other stakeholders to make more evidence-based
decisions for the NRN. CRUSE goes beyond the PCT in several important
ways, modeling denser networks at higher levels of geographic resolution
and additional purposes, including simulated recreational, social,
shopping and personal trips, in addition to trips to work and primary,
secondary and tertiary education locations. The tool provides key
`denominator' information to enable reporting of collision rates per
billion km cycled.

Another key feature of CRUSE is its use of Open Street Map (OSM) data,
which is continuously updated, and used by
\href{https://www.cyclestreets.net/}{Cyclestreets.net} for route
calculation and ``cycle friendliness'' estimates of all links on the
network. CRUSE provides three network types: the `Fastest' network
highlights routes for directness; `Quietest' prioritises cycle
friendliness, and `Balanced' represents a balance between speed and
comfort. Trips are generated based on origin and destination data from
the 2016 Census in combination with modeled demand data to estimate
current and potential future cycling levels at area, route, and network
levels for each county in Ireland.

For investment in cycling to deliver on the huge potential benefits,
cycling must become safer. With growth in the E-bike market, CRUSE will
help inform inter-urban and rural networks to support the transfer of
trips to sustainable modes for longer journeys. Provision of open access
data in a publicly available web tool will address the challenges faced
by National Road Authorities (NRAs) across Ireland in meeting their
sustainability commitments. The experience of developing a deploying
CRUSE are relevant to countries with sustainable transport policies
worldwide.
\end{abstract}

\begin{keyword}
    Cycling \sep Open-source \sep Road Safety \sep 
    Active Travel
\end{keyword}
\end{frontmatter}
    \ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[enhanced, boxrule=0pt, borderline west={3pt}{0pt}{shadecolor}, frame hidden, sharp corners, interior hidden, breakable]}{\end{tcolorbox}}\fi

\newpage{}

@@

Transport is responsible for 23\% of global greenhouse gas emissions,
70\% of which are from road transport. Nearly nearly half of these
emissions (around 10\% of global emissions) are from passenger cars
\citep{jaramillo2022}. The transport system encourages, enables and in
some cases enforces unsustainable lifestyles. Services that are only
accessible by car lock-in car dependency
\citep{gray2001, shergold2012, motte-baumvol2010}.

@@

Uptake of cycling, the main topic of this paper, should be seen in this
broader context of transport decarbonisation \citep{brand2020} and
sustainable mobility \citep{burns2013}. Although cycling uptake appears
on the surface to only relate to the `shift' part of the ASI framework,
closer consideration of the knock-on impacts of cycling uptake shows
that it can also help avoid unnecessary trips \citep{nello-deakin2020}.
Furthermore, highly efficient e-bikes --- which are seeing rapid uptake
in Ireland and other countries --- outperform electric cars, which are
too heavy and expensive. Over-reliance on electric cars could slow the
transition away from car dependency and inadvertently enable ``high
travel lock-in'' \citep{anable2019}. At the European level, the European
Union has a target of reducing greenhouse gas emissions by 55\% by 2030,
compared to 1990 levels, and to achieve `net-zero' by 2050.\footnote{https://eur-lex.europa.eu/legal-content/EN/TXT/?u
  ri=celex\%3A32008L0096}

@@

The main aim of the paper is to present the CRUSE Tool and describe the
methods and datasets on which the estimates of current and future
potential cycling levels are based. We will also highlight key results,
discuss the implications for policy and practice, and outline how
similar tools could be developed in other countries. It is structured as
follows. In Section~\ref{sec-methods}, we outline the methods used to
generate the evidence presented in the CRUSE Tool for Ireland. In
Section~\ref{sec-results}, we present the results of the CRUSE Tool,
including estimates of current cycling levels and future cycling
potential at the national, regional and local levels. In
Section~\ref{sec-discussion}, we discuss the implications of the results
for policy and practice. In Section~\ref{sec-conclusions}, we conclude
the paper.

@@

\begin{itemize}
\tightlist
\item
  Calculation of hundreds-of-thousands of routes, which is beyond the
  terms of service of the free API.
\item
  Making changes to the routing profiles, including allowing routing on
  trunk roads, which are sometimes omitted in the default routing profiles.
\item
  Control over the version of OSM data being used for the routing,
  allowing regular updates to the route networks as OSM data is updated.
\end{itemize}

@@

These map are available on the `Route types' page for each county (see
\href{https://cruse.bike/limerick/route-types}{cruse.bike/limerick/route-types}).
In addition to showing the estimated routes under each scenario, the
page also presents summary statistics on the network:

\begin{itemize}
\item
  On the fastest network, 25\% of the distance cycled occurs in
  non-hostile segments and 6\% in cycle-friendly segments.
\item
  Under the baseline scenario, 55\% of the distance cycled on the
  quietest network occurs in non-hostile segments and 13\% in
  cycle-friendly segments.
\end{itemize}

These statistics can be revealing: in Limerick it suggests that around
half of all cycling activity occurs on network segments that are hostile
(with a high inferred level of traffic stress), even when efforts are
taken to avoid busy segments. For the Fastest network, which may be more
realistic for utility cycling, the proportion of cycling on hostile
segments is even higher at 75\%.

TODO: Add description of route types in Limerick based on local
knowledge and Figure~\ref{fig-route-types}.

@@

\begin{minipage}[t]{\linewidth}

@@

\includegraphics{images/paste-5.png}

@@

\end{minipage}%
\newline
\begin{minipage}[t]{\linewidth}

@@

\includegraphics{images/paste-6.png}

@@

\caption{\label{fig-route-types}Illustration of the Fastest and Quietest
route networks for central and East Limmerick.}

@@

\includegraphics{images/paste-7.png}

@@

\caption{\label{fig-scenarios}Distance-mode share graph for Limerick
used to communicate the different scenarios. Source: county level view
for Limerick in the open access CRUSE Tool at
\url{https://cruse.bike/limerick}.}

@@

The Climate Action Plan scenario is loosely based on the Irish
Government's
\href{https://www.gov.ie/en/publication/6223e-climate-action-plan-2021/}{Climate
Action Plan 2021}, which contains policies for action to achieve a 51\%
reduction in overall greenhouse gas emissions by 2030, on a path towards
net-zero by 2050. For transport, this includes 500,000 extra walking,
cycling and public transport trips per day by 2030. In terms of car
travel, the target is to ``Increas{[}e{]} the proportion of kilometres
driven by passenger electric cars to between 40 and 45\% by 2030, in
addition to a reduction of 10\% in kilometres driven by the remaining
internal combustion engine cars.'' This equates to a 5.5 to 6\%
reduction in total car km driven.

@@

\hypertarget{ebike}{%
\subsection{Ebike}\label{ebike}}

@@

While the CRUSE Tool is not the first national and publicly available
tool for cycling planning, it has some unique features that make it
relevant for other countries, and other national road authorities tasked
with making their transport systems safe.

TODO: outline these features.

The CRUSE Tool is designed to be used by a wide range of stakeholders,
including transport planners, engineers, and other professionals, as
well as non-professionals.

@@

>>>>

\usepackage{amsmath,amssymb}
\usepackage{setspace}
\usepackage{iftex}
\ifPDFTeX
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
  \usepackage{textcomp} % provide euro and other symbols
\else % if luatex or xetex
  \usepackage{unicode-math}
  \defaultfontfeatures{Scale=MatchLowercase}
  \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
\fi
\usepackage{lmodern}
\ifPDFTeX\else  
    % xetex/luatex font selection
\fi
% Use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\IfFileExists{microtype.sty}{% use microtype if available
  \usepackage[]{microtype}
  \UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\makeatletter
\@ifundefined{KOMAClassName}{% if non-KOMA class
  \IfFileExists{parskip.sty}{%
    \usepackage{parskip}
  }{% else
    \setlength{\parindent}{0pt}
    \setlength{\parskip}{6pt plus 2pt minus 1pt}}
}{% if KOMA class
  \KOMAoptions{parskip=half}}
\makeatother
\usepackage{xcolor}
\setlength{\emergencystretch}{3em} % prevent overfull lines
\setcounter{secnumdepth}{5}
% Make \paragraph and \subparagraph free-standing
\ifx\paragraph\undefined\else
  \let\oldparagraph\paragraph
  \renewcommand{\paragraph}[1]{\oldparagraph{#1}\mbox{}}
\fi
\ifx\subparagraph\undefined\else
  \let\oldsubparagraph\subparagraph
  \renewcommand{\subparagraph}[1]{\oldsubparagraph{#1}\mbox{}}
\fi

@@

\makeatletter
\makeatother
\makeatletter
\makeatother
\makeatletter
\@ifpackageloaded{caption}{}{\usepackage{caption}}
\AtBeginDocument{%
\ifdefined\contentsname
  \renewcommand*\contentsname{Table of contents}
\else
  \newcommand\contentsname{Table of contents}
\fi
\ifdefined\listfigurename
  \renewcommand*\listfigurename{List of Figures}
\else
  \newcommand\listfigurename{List of Figures}
\fi
\ifdefined\listtablename
  \renewcommand*\listtablename{List of Tables}
\else
  \newcommand\listtablename{List of Tables}
\fi
\ifdefined\figurename
  \renewcommand*\figurename{Figure}
\else
  \newcommand\figurename{Figure}
\fi
\ifdefined\tablename
  \renewcommand*\tablename{Table}
\else
  \newcommand\tablename{Table}
\fi
}
\@ifpackageloaded{float}{}{\usepackage{float}}
\floatstyle{ruled}
\@ifundefined{c@chapter}{\newfloat{codelisting}{h}{lop}}{\newfloat{codelisting}{h}{lop}[chapter]}
\floatname{codelisting}{Listing}
\newcommand*\listoflistings{\listof{codelisting}{List of Listings}}
\makeatother
\makeatletter
\@ifpackageloaded{caption}{}{\usepackage{caption}}
\@ifpackageloaded{subcaption}{}{\usepackage{subcaption}}
\makeatother
\makeatletter
\@ifpackageloaded{tcolorbox}{}{\usepackage[skins,breakable]{tcolorbox}}
\makeatother
\makeatletter
\@ifundefined{shadecolor}{\definecolor{shadecolor}{rgb}{.97, .97, .97}}
\makeatother
\makeatletter
\makeatother
\makeatletter
\makeatother
\journal{European Transport Research Review}
\ifLuaTeX
  \usepackage{selnolig}  % disable illegal ligatures
\fi
\usepackage[]{natbib}
\bibliographystyle{elsarticle-num}
\IfFileExists{bookmark.sty}{\usepackage{bookmark}}{\usepackage{hyperref}}
\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\urlstyle{same} % disable monospaced font for URLs
\hypersetup{
  pdftitle={CRUSE to Safe Cycling in Ireland},
  pdfauthor={Robin Lovelace; Joey Talbot; Eugeni Vidal-Tortosa; Hussein Mahfouz; Elaine Brick; Peter Wright; Gary O'Tool; Dan Brennan; Suzanne Meade},
  pdfkeywords={Cycling, Open-source, Road Safety, Active Travel},
  colorlinks=true,
  linkcolor={blue},
  filecolor={Maroon},
  citecolor={Blue},
  urlcolor={Blue},
  pdfcreator={LaTeX via pandoc}}

@@

\author[1]{Eugeni Vidal-Tortosa%
%
}

@@

\begin{abstract}
Legislation mandates improved road safety and decarbonisation of
predominantly road-based transport network in Ireland and other
countries. The Cycle Route Uptake and Scenario Estimation (CRUSE) Tool
was commissioned to support cycle network planning, safety
interventions, business case development and project evaluation. Cycling
in Ireland represents only 3\% of total modal share as of the latest
Census (2016), but accounts for 20\% of serious injuries and 7\% of all
fatalities. High resolution datasets on baseline and potential cycling
levels are needed to meet Ireland's climate and safety targets.

The CRUSE Tool is based on open source software and reproducible data
science. The resulting web application is available at
\url{https://cruse.bike/}, enabling planners, engineers, and other
stakeholders to make more evidence-based decisions. CRUSE goes beyond
previous work in several important ways. It models networks at high
levels of geographic resolution, simulates a wide range of trips
including recreational, social, shopping and personal trips, in addition
to trips to work and primary, secondary and tertiary education
locations. The tool provides key `denominator' information to enable
reporting of collision rates per billion km cycled.

Another key feature of CRUSE is its use of frequently-updated
OpenStreetMap data for route calculation and ``cycle friendliness''
estimates of all links on the network. CRUSE provides three network
types: the `Fastest' network highlights routes for directness;
`Quietest' prioritises cycle friendliness, and `Balanced' represents a
balance between speed and comfort. Trips are generated based on origin
and destination data from the 2016 Census in combination with modeled
demand data to estimate current and potential future cycling levels at
area, route, and network levels for each county in Ireland.

For investment in cycling to deliver on the huge potential benefits,
cycling must become safer. With growth in the E-bike market, CRUSE will
help inform inter-urban and rural networks to support the transfer of
trips to sustainable modes for longer journeys. Provision of open access
data in a publicly available web tool will address the challenges faced
by counties across Ireland in meeting their sustainability commitments.
The experience of developing a deploying CRUSE are relevant to countries
with sustainable transport policies worldwide.
\end{abstract}

\begin{keyword}
    Cycling \sep Open-source \sep Road Safety \sep 
    Active Travel
\end{keyword}
\end{frontmatter}
    \ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[sharp corners, enhanced, frame hidden, borderline west={3pt}{0pt}{shadecolor}, breakable, boxrule=0pt, interior hidden]}{\end{tcolorbox}}\fi

\setstretch{2}
\newpage{}

@@

Transport is responsible for 23\% of global greenhouse gas (GHG)
emissions, 70\% of which are from road transport. Nearly nearly half of
these emissions (around 10\% of global emissions) are from passenger
cars \citep{jaramillo2022}. The transport system encourages, enables and
in some cases enforces unsustainable lifestyles. Services that are only
accessible by car lock-in car dependency
\citep{gray2001, shergold2012, motte-baumvol2010}.

@@

Uptake of cycling should be seen in this broader context of transport
decarbonisation \citep{brand2020} and sustainable mobility
\citep{burns2013}. Although cycling uptake appears on the surface to
only relate to the `shift' part of the ASI framework, closer
consideration of the knock-on impacts of cycling uptake shows that it
can also help avoid unnecessary trips \citep{nello-deakin2020}.
Furthermore, highly efficient e-bikes --- which are seeing rapid uptake
in Ireland and other countries --- outperform electric cars, which are
too heavy and expensive. Over-reliance on electric cars could slow the
transition away from car dependency and inadvertently enable ``high
travel lock-in'' \citep{anable2019}. At the European level, the European
Union has a target of reducing GHG emissions by 55\% by 2030, compared
to 1990 levels, and to achieve `net-zero' by 2050.\footnote{https://eur-lex.europa.eu/legal-content/EN/TXT/?u
  ri=celex\%3A32008L0096}

@@

This paper aims to describe the design, features and potential use of
the CRUSE Tool. In Section~\ref{sec-methods}, we outline the methods
used to generate the evidence presented in the CRUSE Tool for Ireland.
In Section~\ref{sec-results}, we present the results of the CRUSE Tool,
including estimates of current cycling levels and future cycling
potential at the national, regional and local levels. In
Section~\ref{sec-discussion}, we discuss the implications of the results
for policy and practice. In Section~\ref{sec-conclusions}, we conclude
the paper.

@@

\begin{itemize}
\tightlist
\item
  Calculation of hundreds-of-thousands of routes, which is beyond the
  terms of service of the free API.
\item
  Making changes to the routing profiles, including allowing routing on
  trunk roads, which are sometimes avoided in the default routing
  profiles.
\item
  Control over the version of OSM data being used for the routing,
  allowing regular updates to the route networks as OSM data is updated.
\end{itemize}

@@

These map are available on the `Route types' page for each county (see
\href{https://cruse.bike/kildare/route-types}{cruse.bike/kildare/route-types}).
In addition to showing the estimated routes under each scenario, the
page also presents summary statistics on the network:

\begin{itemize}
\item
  On the fastest network, 25\% of the distance cycled occurs in
  non-hostile segments and 7\% in cycle-friendly segments.
\item
  Under the baseline scenario, 52\% of the distance cycled on the
  quietest network occurs in non-hostile segments and 15\% in
  cycle-friendly segments.
\end{itemize}

These statistics can be revealing: in Kildare it suggests that around
half of all cycling activity occurs on network segments that are hostile
(with a high inferred level of traffic stress), even when efforts are
taken to avoid busy segments. For the Fastest network, which may be more
realistic for utility cycling, the proportion of cycling on hostile
segments is even higher at 75\%.

@@

\begin{minipage}[t]{0.50\linewidth}

@@

\includegraphics[width=2.60417in,height=\textheight]{images/naas_quietest_godutch.png}

@@

\end{minipage}%
%
\begin{minipage}[t]{0.50\linewidth}

@@

\includegraphics[width=2.60417in,height=\textheight]{images/naas_fastest_godutch.png}

\end{minipage}%
\newline
\begin{minipage}[t]{0.50\linewidth}

\includegraphics[width=2.60417in,height=\textheight]{images/newbridge_quietest_godutch.png}

\end{minipage}%
%
\begin{minipage}[t]{0.50\linewidth}

\includegraphics[width=2.60417in,height=\textheight]{images/newbridge_fastest_godutch.png}

@@

\caption{\label{fig-route-types}Illustration of the Quietest (left) and
Fastest (right) route networks for Naas (top) and Newbridge (bottom).}

@@

\includegraphics{images/scenarios_kildare.png}

@@

\caption{\label{fig-scenarios}Distance-mode share graph for Kildare used
to communicate the different scenarios. Source: county level view for
Kildare in the open access CRUSE Tool at
\url{https://cruse.bike/kildare}.}

@@

The Climate Action Plan scenario is loosely based on the Irish
Government's
\href{https://www.gov.ie/en/publication/6223e-climate-action-plan-2021/}{Climate
Action Plan 2021}, which contains policies for action to achieve a 51\%
reduction in overall GHG emissions by 2030, on a path towards net-zero
by 2050. For transport, this includes 500,000 extra walking, cycling and
public transport trips per day by 2030. In terms of car travel, the
target is to ``Increas{[}e{]} the proportion of kilometres driven by
passenger electric cars to between 40 and 45\% by 2030, in addition to a
reduction of 10\% in kilometres driven by the remaining internal
combustion engine cars.'' This equates to a 5.5 to 6\% reduction in
total car km driven.

@@

\hypertarget{ebike}{%
\subsubsection{Ebike}\label{ebike}}

@@

While the CRUSE Tool is not the first national and publicly available
tool for cycling planning, it has some key features that make it
relevant for other countries, and other national road authorities tasked
with making their transport systems safe:

\begin{itemize}
\tightlist
\item
  The results are \emph{open access} meaning that any stakeholder in the
  planning system can access the evidence. This will help to democratise
  the transport planning process and make wider conversations about
  transport planning more evidence-based and less polarised
  \citep{lovelace2020}, something that is particularly relevant given
  the potentially polarising nature of pro-cycling interventions
  \citep{wild2017}.
\item
  The results are fully reproducible (code to be released pending
  sign-off by TII's IT team), preventing `cloud lock in' to a
  potentially monopolistic consultancy, and encouraging input from the
  wider open source community \citep{lovelace2021, dhir2017}.
\item
  By covering a wide range of trip purposes --- not just travel to work
  \citep{lovelace2017, heinen2010} and travel school \citep{goodman2019}
  as covered in previously published research on open access national
  cycle netework planning tools --- the results capture a higher
  proportion of cycling potential including key rural trips which are
  often under-represented in models of active travel.
\end{itemize}

@@

The CRUSE Tool is an open access web application for strategic cycle
network planning and prioritisation of road safety interventions across
Ireland. Building on previous work, it provides a nationally consistent
evidence base, providing valuable insights to planners and other
stakeholders in the transport planning process, at national and local
levels. Because the results are available at the route segment level,
the tool can be used to identify `weak links' in the cycle network, and
to prioritise investment in cycling infrastructure. Furthermore, the
provision of the evidence in a free and publicly available website,
hosted at \href{https://cruse.bike}{cruse.bike}, means that it can be
used by anyone, encouraging wider participation and more evidence-based
debate about transport planning.

A key feature of the project methodologically is its calculation of
current and future potential not only for travel to work and travel to
school, but also for other trip purposes, including recreational trips.
This required the development of spatial interaction models and
estimation of the relative attractiveness of different destinations for
different trip purposes, an area of active research where new
developments could be incorporated into the tool in future
\citep{hasova2022}.

The CRUSE Tool and the underlying research and methods are not without
limitations, suggesting future areas of research, data collection needs,
and policy application. The route network level results have not been
validated to the extent we had planned at the outset of the project. We
compared route network level results with `ground truth' data from cycle
counter datasets across Dublin to test different network types and to
ensure that additional trip types added to work and educational trips
increased the quality of fit under the baseline scenario. However, the
size of the counter network was insufficient to provide an opportunity
for robust evaluation of model performance, suggesting a combined
program of new count data collection and analysis should be a priority
for future work, with findings feeding directly into work to improve the
outputs of the the CRUSE Tool.

More broadly, the tool's outputs are limited to just one more of active
travel (cycling), ignoring walking and wheeling, including wheelchair
use and a range of wheeled devices such as scooters, bike trailers and
e-cargo bikes that can be used to escort children to school. Sustainable
transport policies should plan for walking, wheeling and cycling, and
there are many co-benefits of broadly defined active travel
interventions that benefit all active modes, such as measures to reduce
heavy motor traffic speeds and volumes in areas and along corridors with
high active travel potential. This raises the question of whether other
active modes should be incorporated into the results using the OD-based
approach outlined in this paper, or whether different modelling
approaches are needed to properly capture the shorter distance trips
typically made by walking \citep{cooper2018}. Furthermore, the estimates
of cycling presented in the CRUSE Tool omit multi-modal trips including
public transport, and omit trip chaining, due to the need to capture a
large portion of cycling potential within the resource constraints of
the project.

The CRUSE Tool is already used in practice to support more ambitious and
data-driven planning for safe cycling routes in multiple counties across
Ireland. We hope that the underlying approach, and the publicly
available evidence provided in the CRUSE Tool itself, provide a basis
for future reproducible research, open source software development into
strategic cycle network planning tools in Ireland and other countries.
Active travel represents a win-win-win for physical activity,
environment, and local economic opportunities. In combination with
broader sustainable mobility measures and policies to reduce motor
traffic speeds and volumes, tools such as CRUSE can support the fast and
fair decarbonisation of transport systems worldwide.
